---
title: "GitHub Actions for KWB-R Packages"
author: "Michael Rustler, Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GitHub Actions for KWB-R Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig_caption = FALSE,
  comment = "#>",
  eval = FALSE
)
```

`kwb.pkgbuild` ships a curated set of [GitHub
Actions](https://github.com/features/actions) workflows that we use across
all KWB-R packages. The workflows live as templates in
`inst/templates/ci_github-actions/` (default workflows) and
`inst/templates/ci_github-actions-claude/` (optional Claude Code
integration), and are installed into a package's `.github/workflows/`
directory by `use_ghactions()` / `use_pkg()`.

This vignette explains what the workflows do, how to add or update them, and
how to enable the optional Claude Code integration.

# 1 Default workflows

`kwb.pkgbuild::use_ghactions()` (also called from `kwb.pkgbuild::use_pkg()`)
copies four workflows into `.github/workflows/`:

| File                  | Purpose                                                                                                  |
|:----------------------|:---------------------------------------------------------------------------------------------------------|
| `R-CMD-check.yaml`    | Cross-platform `R CMD check` matrix (macOS, Ubuntu, Windows × `release`, `devel`, `oldrel-1`).           |
| `pkgdown.yaml`        | Builds the `pkgdown` site and deploys it to the `gh-pages` branch on push to `main`/`master`/`dev`.      |
| `test-coverage.yaml`  | Runs `covr::package_coverage()` and uploads the report to [codecov.io](https://codecov.io).              |
| `pr-commands.yaml`    | Reacts to `/document` and `/style` comments on pull requests by running `roxygen2` / `styler` for you.    |

All four use the official [`r-lib/actions`](https://github.com/r-lib/actions)
v2 building blocks, which means dependency installation and caching are
handled by `r-lib/actions/setup-r-dependencies@v2`.

To install or refresh the default workflows in a package, run:

```{r}
# from the package root
kwb.pkgbuild::use_ghactions()
```

# 2 Optional: Claude Code workflows

If you want [Claude Code](https://docs.claude.com/en/docs/claude-code/) to
review your pull requests and answer `@claude` mentions in issues / PR
comments, set `claude = TRUE` when bootstrapping a new package:

```{r}
kwb.pkgbuild::use_pkg(
  author      = author,
  pkg         = description,
  version     = "0.0.0.9000",
  stage       = "experimental",
  claude      = TRUE
)
```

Or add the workflows to an existing package after the fact:

```{r}
kwb.pkgbuild::use_ghactions_claude()
```

This installs two additional workflow files:

| File                       | Trigger                                                                                                     |
|:---------------------------|:------------------------------------------------------------------------------------------------------------|
| `claude.yaml`              | `issue_comment`, `pull_request_review_comment`, `pull_request_review`, `issues` — reacts to `@claude` mentions. |
| `claude-code-review.yaml`  | `pull_request` (`opened`, `synchronize`) — automatic PR review.                                              |

## 2.1 Required repository secret

Both workflows authenticate via a GitHub Actions secret named
`CLAUDE_CODE_OAUTH_TOKEN`. To create it:

1. Open the repository on GitHub and go to
   **Settings → Secrets and variables → Actions → New repository secret**.
2. Set **Name** to `CLAUDE_CODE_OAUTH_TOKEN` and paste your Claude Code OAuth
   token as **Value**.

Without this secret the Claude jobs will fail at the
`anthropics/claude-code-action@v1` step.

## 2.2 What Claude does

`claude.yaml` runs whenever someone mentions `@claude` in an issue body /
title, an issue comment, a PR review or a PR review comment. Claude then
opens a session with read access to issues / pull requests, plans, and
posts a reply.

`claude-code-review.yaml` runs on every newly opened or updated pull
request. It reviews the diff with focus on:

- code quality and R package best practices,
- potential bugs or logic issues,
- test coverage,
- `roxygen2` documentation completeness,
- consistency with the existing codebase.

If your repository contains a `CLAUDE.md` file, Claude will use it as
project-specific guidance.

# 3 Updating workflows in existing packages

The workflows in `inst/templates/ci_github-actions/` are the source of truth
for KWB-R packages. When upstream actions are updated (e.g. a new version of
`r-lib/actions`) we bump the templates here in `kwb.pkgbuild` and re-run
`kwb.pkgbuild::use_ghactions()` in each package to refresh the local
workflows.

To check whether a package is using an outdated version, simply diff its
`.github/workflows/` against the templates in the latest installed
`kwb.pkgbuild`:

```{r}
template_dir <- system.file(
  "templates/ci_github-actions",
  package = "kwb.pkgbuild"
)

# from the package root
list.files(".github/workflows", full.names = TRUE)
list.files(template_dir,         full.names = TRUE)
```

# 4 Troubleshooting

- **The pkgdown workflow doesn't deploy.** Make sure the `gh-pages` branch
  exists and that **Settings → Pages → Source** points to it. The first
  time a package is set up, run `kwb.pkgbuild::use_autopkgdown()` to create
  the orphan branch.

- **`R CMD check` fails on `oldrel-1`.** Either pin a `Depends: R (>= ...)`
  in `DESCRIPTION` to the lowest R version you actually want to support, or
  remove `oldrel-1` from the matrix in `R-CMD-check.yaml`.

- **Claude jobs are skipped.** The `if:` clause requires `@claude` to appear
  in the comment / review / issue body. The Claude review workflow only
  runs on `pull_request`, not on `issue_comment` events.

- **Codecov upload fails on forks.** Configure the `CODECOV_TOKEN`
  repository secret. The workflow has `fail_ci_if_error` set conditionally
  so PRs from forks won't break the build.
